---

author: "Michael Campbell"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(plyr)
library(dplyr)
library(knitr)
library(lubridate)
library(plm)
library(broom)
library(lfe)
library(sandwich)
library(tidyr)
library(data.table)
library(MatchIt)
library(ggplot2)


### Setwd
setwd("Y:/<insert path>")

### Import Data
August_treat <- fread("aug_16_event_TREATMENT_negFixed.csv", data.table = FALSE)
July_treat <- fread("july_13_event_TREATMENT_negFixed.csv", data.table = FALSE)
August_control <- fread("aug_16_event_CONTROL_negFixed.csv", data.table = FALSE)
July_control <- fread("july_13_event_CONTROL_negFixed.csv", data.table = FALSE)
Customer_Data <- fread("Customer Data.csv", data.table = FALSE)

```

The number of unique observations in August treatment is ` r length(unique(August_treat$custID))`. (4519)
The number of unique observations in Augst control is ` r length(unique(August_control$custID))`. (4448)
The number of unique observations in July treatment is ` r length(unique(July_treatment$custID))`. (4519)
The number of unique observations in July control is ` r length(unique(July_control$custID))`. (4448)
```{r}
Treat_Merge <- rbind(August_treat, July_treat)
Control_Merge <- rbind(August_control, July_control)



### Combining Modified datasets

Master_dataset <- rbind(Control_Merge, Treat_Merge)

# matching custID types
Master_dataset$custID <- as.numeric(Master_dataset$custID)

Master_dataset <- left_join(Master_dataset, Customer_Data, by=c("custID"))

### Drop Redundant Columns

#Master_dataset <- subset(Master_dataset, select = -c(V1.x, X, V1.y))
```


```{r}
### Create working dataset

data <- Master_dataset

### Convert timestamp to usable format and adjust time to hour end by subtracting 3600 seconds. 00:00:00 becomes
### 23:00:00

data$timestamp <- strptime(data$timestamp, format = "%d%b%Y:%H:%M:%S") - rep(3600, length(data$timestamp))
data$minute <- minute(data$timestamp)
data$hour <- hour(data$timestamp)
data$day <- day(data$timestamp)
data$month <- month(data$timestamp)

### Showing occurrence of non-hourly readings
hist(subset(data, minute!=0)$minute,
xlab = "Minute of Occurence",
main = "Occurrence of Non-Hourly Readings")
axis(1, at = seq(0, 60, by = 5))
hist(subset(data, minute!=0)$minute, breaks = unique(data$minute),
     xlab = "Minute of Occurence",
     main = "Occurence of Non-Hourly Readings")
axis(1, at = seq(0, 60, by = 5))

## Sort by id and then by time
sortvect <- order(data[, "custID"], data[,"timestamp"])
data <- data[sortvect,]

### drop the non-hourly observations readings. before: 2,247,397. Unique custID: 8,969.
data <- subset(data, minute == 0)
### After 2,230,575. Unique custID: 8,969. Total dropped: 16,822

## Create UsageVariable
data <- data %>% 
      group_by(custID, month) %>% 
      mutate(usage = read_value - lag(read_value, default=first(read_value))) %>%
      ungroup()



### Subset months to drop 1st hour zero usage period
# 1,115,569 observations before drop
aug_sub <- subset(data, month == 8)
aug_sub <- aug_sub[duplicated(aug_sub$custID),]
# 1,106,602 observations after drop.
aug_sub <- aug_sub %>% 
    group_by(custID, day) %>%
    filter(length(hour) <= 24 & day != 12) %>%
    ungroup()
# 953,743 observations after drop with 8,653 unique custIDs.

july_sub <- subset(data, month == 7)
# 1,115,006 observations
july_sub <- july_sub[duplicated(july_sub$custID),]

# 1,106,041 observations after drop and 8,965 unique IDs.
july_sub <- july_sub %>% 
    group_by(custID, day) %>%
    filter(length(hour) <= 24 & day != 9) %>%
    ungroup()
# 952,079 observations after drop and 8,647 unique IDs.

## Recombine into working dataset

data <- rbind(july_sub, aug_sub) #total of 1,905,822 observations and 8,671 unique IDs.


### Data Exploration and Analysis

## na count after dropping first consumption period
# before dropping nas: 1,905,822 and 8,671 unique IDs.
na_count <-sapply(data, function(y) sum(length(which(is.na(y)))))
print(na_count)
data <- na.omit(data, "usage")
# After dropping: 1,901,349 observations and 8,613 unique IDs.
# Keeping custIDs with negative readings, but dropping those observations.

negative <- subset(data, usage < 0) #1,085 observations across 379 observations
positive <- subset(data, usage >= 0) #1,900,264 observations

    
  
```

Graph for consumption.
```{r}

  

aug_event <- subset(positive, day == "16" & month == "8")
july_event <- subset(positive, day == "13" & month == "7")


## create usage graph data

summary <- spread(aggregate(usage ~ Treatment + hour, aug_event, FUN = "mean"), Treatment, usage)
colnames(summary) <- c("Hours", "Control", "Treatment")

ggplot(summary, mapping = aes(Hours, y = Consumption, color = Variables)) + 
  geom_line(aes(y = Control, col = "Control")) + 
  geom_line(aes(y = Treatment, col = "Treatment"))

```


```{r}
## Climate Zones
clim.obs <- ddply(data, ~Climate, summarise, distinct_czones=length(unique(custID)))

## Rural vs Urban
metro.obs <- ddply(data, ~Metro, summarise, distinct_metro=length(unique(custID)))

## Tenant Status

tenant.obs <- ddply(data, ~NewTenant, summarise, distinct_tenant=length(unique(custID)))

## Zipcode distribution

zip.obs <- ddply(data, ~Zip, summarise, distinct_zip=length(unique(custID)))
zip.obs <- zip.obs[order(zip.obs$distinct_zip, decreasing =TRUE),]

## consumption graph

aug_event <- subset(data, day == "16" & month == "8")
july_event <- subset(data, day == "13" & month == "7")

## drop NA for graph
aug_event <- na.omit(aug_event, "usage")
july_event <- na.omit(july_event, "usage")

## create usage graph data

#summary <- spread(aggregate(usage ~ Treatment + hour, aug_event, FUN = "mean"), Treatment, usage)
#colnames(summary) <- c("Hours", "Control", "Treatment")
#ggplot(summary, mapping = aes(Hours, y = Consumption, color = Variables)) + 
  #geom_line(aes(y = Control, col = "Control")) + 
  #geom_line(aes(y = Treatment, col = "Treatment"))
``` 
